%Abstract
With the development of web technology, web applications, or web apps for short, have had a profound effect on the way people get informations and communicate with one another. As web apps can be developed to have similar functionality to the native apps and accessed on a range of different platforms, recent trends have shown much interest in the development of web apps, which are normally built using HTML, CSS and JavaScript. This report presents the development of The Organ Web App which aims to share the historically important heritage of the Henry Willis pipe organ with as wide an audience as possible.

In particular, this report presents the findings of a five months advanced project placement on the study of organ, undertaken at the Union Chapel. A web app that introduces different aspects of the organ is presented, which also features a small collection of audio samples from the organ in order to let the user press the keyboard and hear different timbres. The app can be used in a creative interactive way to make the Henry Willis organ more accessible to the local community and beyond, especially for people who have never been to the Union Chapel.

This report outlines the following: how digital technologies have engendered new music application; notices for web application design; introductions of the main web technologies applied in the app development; a detailed explanation of the implementation process. The app is evaluated by online questionnaire: positive and negative aspects of the design are identified. The Organ Web App as a whole serves as a tool to provide users with a comprehensive exploration of the Henry Willis pipe organ.


%This report outlines the following: how digital technologies have engendered new music application; notices for web application design; introductions of the main web technologies applied in the implementation process. The Organ Web App is evaluated by online questionnaire: positive and negative aspects of the design are identified, as well as aspects that would be improved in the future work. The app as a whole serves as a tool to provide users with a comprehensive exploration of the Henry Willis pipe organ.







